Weighted Model Counting Without Parameter Variables

### 2021-02-20 doc/paper3/paper.tex
New template

\documentclass{article}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[UKenglish]{babel}
\usepackage[capitalise]{cleveref}
\usepackage{fullpage}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{isodate}
\usepackage{mathtools}
\usepackage[backgroundcolor=lightgray]{todonotes}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

@@

\title{Weighted Model Counting Without Parameter Variables}

@@

\maketitle

@@

% NOTE: ...that we will need.
% NOTE: The proof immediately comes from the definition (of operations).
\begin{proposition}[Basic properties] \label{prop:basic}
  For any propositional formulas $\phi$ and $\psi$, and $a, b, c, d \in
  \mathbb{R}$,
  \begin{itemize}
  \item $[\phi]^a_b = [\neg \phi]^b_a$;
  \item $c + [\phi]^a_b = [\phi]^{a+c}_{b+c}$;
  \item $c \cdot [\phi]^a_b = [\phi]^{ac}_{bc}$;
  \item $[\phi]^a_b \cdot [\phi]^c_d = [\phi]^{ac}_{bd}$;
  \item $[\phi]^1_0 \cdot [\psi]_0^1 = [\phi \land \psi]_0^1$.
  \end{itemize}
\end{proposition}

@@

\paragraph{Notes.}
\begin{itemize}
\item Apparently, the \textsc{DPMC} paper already shows that taking the first
  offered decomposition tree is best.
\item It is already well-known that WMC is FPT.
\item Weights on literals other than the positive literals that correspond to
  variables in $X_P$ are redundant as they either are equal to one or duplicate
  an already-defined weight.
\item Mention that formulas, clauses, and models are all treated as sets.
\end{itemize}

@@

\begin{algorithm}
  \caption{WMC instance transformation}
  \label{alg:transformation}
  \SetKwFunction{rename}{rename}
  \SetKwProg{Fn}{Function}{:}{}
  \KwData{an (old-format) WMC instance $(\phi, X_I, X_P, w)$}
  \KwResult{a (new-format) WMC instance $(F, X, \omega)$}
  $F \gets \emptyset$\;
  $\omega \gets 1$\;
  let $d\colon X_P \to \mathbb{N}$ be defined as $p \mapsto |\{ o \in X_P \mid o
  \le p \}|$\;
  \ForEach{clause $c \in \phi$\label{line:foreach1start}}{
    \uIf{$c \cap X_P = \{ p \}$ for some $p$ \textnormal{\textbf{and}} $w(p) \ne
      1$}{
      \uIf{$|c| = 1$}{$\omega \gets \omega \times w(p)$\;}
      \Else{
        $F \gets F \cup \left\{ \left[ \bigwedge_{l \in c \setminus
              \{p\}} \neg l \right]^{w(p)}_1 \right\}$\;
      }
    }
    \ElseIf{$\{p \mid \neg p \in c\} \cap X_P  = \emptyset$}{
      $F \gets F \cup \{ [c]^1_0 \}$\; \label{line:foreach1end}
    }
  }
  \ForEach{indicator variable $v \in X_I$}{
    \If{$\{[v]_1^p, [\neg v]_1^q\} \subseteq F$ for some $p$ and $q$}{
      $F \gets F \setminus \{ [v]_1^p, [\neg v]_1^q \} \cup \{ [v]_q^p
      \}$\;
    }
  }
  replace every variable $v$ in $F$ with $\rename{$v$}$\;
  \Return{$(F, X_I, \omega)$}\;
  \Fn{\rename{$v$}}{
    $S \gets \{u \in X_P \mid u \le v\}$\;
    \lIf{$S = \emptyset$}{\Return{$v$}}
    \Return{$v - d(\max S)$}\;
  }
\end{algorithm}

@@

\begin{theorem}[Correctness]
  \Cref{alg:transformation}, when given an old WMC instance $(\phi, X_I, X_P,
  w)$, returns a new WMC instance with the same answer, provided the following
  conditions are satisfied:
  \begin{enumerate}
  \item for all indicator variables $i \in X_I$, $w(i) = w(\neg i) = 1$,
  \item and either
    \begin{enumerate}
    \item for all parameter variables $p \in X_P$, there is a non-empty family
      of literals $(l_i)_{i=1}^n$ such that \label{condition:d02}
      \begin{enumerate}
      \item $w(\neg p) = 1$,
      \item $l_i \in X_I$ or $\neg l_i \in X_I$ for all $i = 1, \dots, n$,
        \label{condition:equivalence1}
      \item and $\{ c \in \phi \mid p \in c \text{ or } \neg p \in c \} =
        \left\{p \lor \bigvee_{i=1}^n \neg l_i \right\} \cup \{ l_i \lor \neg p
        \mid i = 1, \dots, n \}$; \label{condition:equivalence2}
      \end{enumerate}
    \item or for all parameter variables $p \in X_P$, \label{condition:bklm16}
      \begin{enumerate}
      \item $w(p) + w(\neg p) = 1$,
      \item for any clause $c \in \phi$, $|c \cap X_P| \le 1$,
      \item there is no clause $c \in \phi$ such that $\neg p \in c$,
      \item if $\{ p \} \in \phi$, then there is no clause $c \in \phi$ such
        that $c \ne \{ p \}$ and $p \in c$, \label{cond:just_parameter}
        % 'prime implicants are logically disjoint'
      \item and for any $c, d \in \phi$ such that $c \ne d$, $p \in c$ and $p
        \in d$, $\bigwedge_{l \in c \setminus \{ p \}} \neg l \land \bigwedge_{l
          \in d \setminus \{ p \}} \neg l$ is false. \label{cond:disjoint}
      \end{enumerate}
    \end{enumerate}
  \end{enumerate}
\end{theorem} % TODO: add parentheses around products
% TODO: maybe give a name to (2) since it's wrong to call it an equation
% TODO: check if each condition is actually used. Maybe turn this into a
% paragraph that gives an overview of the proof.
\begin{proof}
  By \cref{prop:equivalence},
  \begin{equation}
    \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
          x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
      1\right) \label{eq:new_wmc}
  \end{equation}
  is a new WMC instance with the same answer as the given old WMC instance. By
  \cref{def:new_wmc}, its answer is
  \begin{equation} \label{eq:answer}
    \left(\exists_{X_I \cup X_P} \prod_{c \in \phi} [c]_0^1 \prod_{x \in X_I
        \cup X_P} [x]_{w(\neg x)}^{w(x)} \right)(\emptyset)
  \end{equation}
  Since both \cref{condition:d02,condition:bklm16} ensure that each clause in
  $\phi$ has at most one parameter variable, we can partition $\phi$ into
  $\phi_* \coloneqq \{c \in \phi \mid \mathtt{Vars}(c) \cap X_P = \emptyset \}$
  and $\phi_p \coloneqq \{ c \in \phi \mid \mathtt{Vars}(c) \cap X_P = \{ p \}
  \}$ for all $p \in X_P$. We can then use \cref{thm:early} to reorder
  \cref{eq:answer} into
  \[
    \left(\exists_{X_I} \left( \prod_{x \in X_I} [x]_{w(\neg x)}^{w(x)} \right)
      \left( \prod_{c \in \phi_*} [c]_0^1 \right) \prod_{p \in X_P} \exists_p
      [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 \right)(\emptyset).
  \]

@@

  The only remaining case is when $w(p) \ne 1$ and $\{ p \} \not \in \phi_p$.
  Then $\omega = 1$, and
  \[
    F_p = \left\{ \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
        l\right]_1^{w(p)} \;\middle|\; c \in \phi_p \right\},
  \]
  so need to show that
  \[
    \prod_{c \in \phi_p} \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
      l\right]_1^{w(p)} = \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p}
    [c]_0^1.
  \]
  We can rearrange the right-hand-side as
  \begin{align*}
    \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[\bigwedge_{c \in \phi_p} c\right]_0^1 = \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 = w(p) + \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^{1-w(p)} \\
                                                               &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \neg \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \neg(c \setminus \{ p \}) \right]_1^{w(p)} \\
                                                               &= \left[ \bigvee_{c \in \phi_p} \neg \bigvee_{l \in c \setminus \{ p \}} l \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}
  \end{align*}
   by \cref{prop:basic,def:projection}. By \cref{cond:disjoint}, $\bigwedge_{l
     \in c \setminus \{ p \}} \neg l$ can be true for at most one $c \in
   \phi_p$, and so
   \[
     \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l
     \right]_1^{w(p)} = \prod_{c \in \phi_p} \left[ \bigwedge_{l \in c \setminus
         \{ p \}} \neg l \right]_1^{w(p)}
   \]
   which is exactly what we needed to show.
\end{proof}
% TODO: now consider the second for loop
% TODO: condition 1 is necessary in both cases because we're ignore the weights
% of indicator variables
% NOTE: benefits of having the proof in the paper:
% 1) it puts all the encodings on a common ground
% 2) it illustrates the convenience of our notation for reasoning about (certain
% types of) pseudo-Boolean functions (stress this!!!)
% 3) it's too big and too important to be left for the appendix
% NOTE: we don't make a distinction between a real number and a (PB) function
% that always returns that number.
% NOTE: projection is assumed to have lowest precedence (i.e., \exists_x fg =
% \exists_x (fg) and not (\exists_x f)g).
% TODO: need another basic property: (fg)|_{x=i} = f|_{x=i} * g|_{x=i} for i=0, 1.
% TODO: \times -> \cdot

\paragraph{Notes.}
\begin{itemize}
\item This just says that $p$ is equivalent to a conjunction.
\item The first group of conditions applies to \texttt{d02}, while the second
  group applies to \texttt{bklm16}.
\item For \texttt{cd05} and \texttt{cd06}, condition $2bi$ should be replaced
  with $w(\neg p) = 1$.
\end{itemize}

\section{Parameterised Complexity of \textsc{DPMC}}

\paragraph{Notes.}
\begin{itemize}
\item Summary of results
  \begin{itemize}
  \item We establish \textsc{DPMC} inference as fixed-parameter tractable.
  \item We experimentally show that \textsc{DPMC} is best on low-to-moderate
    treewidth instances, and cd06+c2d overtakes on higher treewidth instances.
  \end{itemize}
\item By \textsc{DPMC}, we always mean DMC+lg.
\item The constant is inspired by the \texttt{bklm16} encoding.
\end{itemize}

\paragraph{TODO}
\begin{itemize}
\item Do I need to formally consider extending a pseudo-Boolean function to a
  bigger domain?
\item Introduce all the notation surrounding graphs ($\mathcal{V}$,
  $\mathcal{E}$ $\mathcal{L}$, $\mathcal{C}$, etc.) (and be consistent with its
  usage) and formulas (variable, literal, clause, CNF, etc.)
\item Use \cref{thm:connection} to show the connection between primal treewidth
  and the width of the PJT (linking the complexity of BN inference with the DPMC
  complexity).
\item Come up with better names for old/new WMC instances.
\item Define:
  \begin{itemize}
  \item scalar operations on ADDs,
  \item an ADD as a DAG,
  \item What does it mean for an ADD to `have' variables? Maybe refer to
    pseudo-Boolean function sensitivity.
  \item primal graph of a CNF formula (a.k.a.
    Gaifman/(variable) interaction/connectivity/clique/representing graph),
  \item Bayesian network (I already have a definition of these last two),
  \item moralisation of a Bayesian network (or of any DAG),
  \item projection of a set of variables.
  \end{itemize}
\end{itemize}

\begin{lemma} \label{lemma:add_size}
  An ADD with $n$ variables has $\mathcal{O}(2^n)$ nodes.
\end{lemma}
\begin{proof}
  An ADD with $n$ variables can be, at most, a complete binary tree of height
  $n+1$ (as measured by the number of vertices). It would then have $2^0 + 2^1 +
  \dots + 2^n = 2^{n+1} = \mathcal{O}(2^n)$ nodes.
\end{proof}

\begin{lemma} \label{lemma:clause_time}
  Let $\phi$ be a conjunction of $n$ literals. Then the ADD representation of
  $[\phi]^p_q$ can be constructed in $\mathcal{O}(2^n)$ time for any $p, q \in
  \mathbb{R}$ such that $p \ne q$.
\end{lemma}
\begin{proof}
  The ADD representation of $\phi$ itself can be constructed with a sequence of
  $n-1$ calls to \texttt{apply} with one of the two operands in each call always
  a literal. The number of variables in the other operand then follows the
  sequence $1, 2, 3, \dots, n-1$. By \cref{lemma:add_size}, the numbers of nodes
  in the ADD representations of these operands is then $\mathcal{O}(2^1),
  \mathcal{O}(2^2), \dots, \mathcal{O}(2^{n-1})$. Since one of the operands is
  of constant size, the overall time complexity of all calls to \texttt{apply}
  is then
  \[
    \mathcal{O}(2^1) + \mathcal{O}(2^2) + \dots + \mathcal{O}(2^{n-1}) =
    \mathcal{O}(2^n).
  \]

  Let $\alpha$ be the ADD representation of $\phi$. Then the ADD representation
  of $[\phi]_q^p$ is $(p-q)\alpha + q$. As scalar operations can obviously be
  implemented in linear time, the overall complexity remains $\mathcal{O}(2^n)$.
\end{proof}

\begin{theorem}[\cite{DBLP:conf/ecai/KwisthoutBG10}, rephrased]
  BN Inference (for all algorithms that accept arbitrary instances) has a lower
  bound that's linear in the size of the BN and exponential in the treewidth of
  its moralisation (provide the exact formula).
\end{theorem}

\begin{definition}[\cite{DBLP:conf/cp/DudekPV20}, with some changes]
  Let $X$ be a set of Boolean variables, and $F$ be a set of pseudo-Boolean
  functions $2^X \to \mathbb{R}$. A \emph{project-join tree} (PJT) of $F$ is a
  tuple $(T, r, \gamma, \pi)$ where:
  \begin{itemize}
  \item $T$ is a tree with root $r \in \mathcal{V}(T)$,
  \item $\gamma\colon \mathcal{L}(T) \to F$ is a bijection between the leaves of
    $T$ and the pseudo-Boolean functions in $F$, and
  \item $\pi\colon \mathcal{V}(T) \setminus \mathcal{L}(T) \to 2^X$ is a
    labelling function on internal nodes.
  \end{itemize}
  Moreover, $(T, r, \gamma, \pi)$ must satisfy the following two properties:
  \begin{enumerate}
  \item $\{\pi(n) : n \in \mathcal{V}(T) \setminus \mathcal{L}(T)\}$ is a
    partition of $X$, and
  \item for each internal node $n \in \mathcal{V}(T) \setminus \mathcal{L}(T)$,
    variable $x \in \pi(n)$, and `clause' $c \in F$ such that $x$
    appears in $c$, the leaf node $\gamma^{-1}(c)$ must be a descendant of $n$
    in $T$.
  \end{enumerate}
\end{definition}

\begin{definition}[\cite{DBLP:conf/cp/DudekPV20}] \label{def:size}
  \[
    \mathtt{Vars}(n) \coloneqq
    \begin{cases}
      \mathtt{Vars}(\gamma(n)) & \text{if } n \in \mathcal{L}(T) \\
      \left( \bigcup_{o \in \mathcal{C}(n)} \mathtt{Vars}(o) \right) \setminus
      \pi(n) & \text{if } n \not\in \mathcal{L}(T).
    \end{cases}
  \]
  \[
    \mathtt{size}(n) \coloneqq
    \begin{cases}
      |\mathtt{Vars}(n)| & \text{if } n \in \mathcal{L}(T) \\
      |\mathtt{Vars}(n) \cup \pi(n)| & \text{if } n \not\in \mathcal{L}(T).
    \end{cases}
  \]
  (Note that $\mathtt{size}(n)$ is the number of variables that can appear
  during the computation of $\delta(n)$ (excluding recursive calls).)
  The \emph{width} of a PJT $(T, r, \gamma, \pi)$ is
  $\mathtt{width}(T) \coloneqq \max_{n \in \mathcal{V}(T)} \mathtt{size}(n)$.
\end{definition} % TODO: don't actually need this much notation.

\begin{theorem}[\cite{DBLP:conf/cp/DudekPV20}, rephrased, `ADD width is equal
  to the treewidth of the primal graph']
  Given a CNF formula $\phi$ with a tree decomposition of its primal graph of
  width $w$, Algorithm~2 (from \cite{DBLP:conf/cp/DudekPV20}) returns a
  PJT of $\phi$ of width at most $w+1$.
\end{theorem}

\begin{definition}[\cite{DBLP:journals/jct/RobertsonS84}, rephrased]
  A \emph{tree decomposition} of a graph $G$ is a pair $(T, \chi)$, where $T$ is
  a tree and $\chi\colon \mathcal{V}(T) \to 2^{\mathcal{V}(G)}$ is a labelling
  function, with the following properties:
  \begin{itemize}
  \item $\bigcup_{t \in \mathcal{V}(T)} \chi(t) = \mathcal{V}(G)$;
  \item for every edge $e \in \mathcal{E}(G)$, there exists $t \in
    \mathcal{V}(T)$ such that $e$ has both endpoints in $\chi(t)$;
  \item for all $t, t', t'' \in \mathcal{V}(T)$, if $t'$ is on the path between
    $t$ and $t''$, then $\chi(t) \cap \chi(t'') \subseteq \chi(t')$.
  \end{itemize}
  The \emph{width} of tree decomposition $(T, \chi)$ is $\max_{t \in
    \mathcal{V}(T)} |\chi(t)| - 1)$. The \emph{treewidth} of graph $G$ is the
  smallest $w$ such that $G$ has a tree decomposition of width $w$.
\end{definition}

\begin{lemma} \label{lemma:basic_ops}
  Let $f,g\colon 2^X \to \mathbb{R}$ be pseudo-Boolean functions represented by
  ADDs with $n$ and $m$ nodes, respectively, and $x \in X$. Then $f+g$ and $fg$
  can be computed in $\mathcal{O}(mn)$ time, and $\exists_xf$ can be computed in
  $\mathcal{O}(n^2)$ time.
\end{lemma}
\begin{proof}
  Addition and multiplication are implemented by \texttt{apply} algorithm which
  takes $\mathcal{O}(mn)$ time \cite{DBLP:journals/tc/Bryant86}. Projection
  consists of two restrictions and an addition by \cref{def:projection}. The
  computational complexity of restriction is dominated by the reduction
  operation that transforms a decision diagram into a minimal canonical form
  \cite{DBLP:journals/tc/Bryant86}. While the original reduction algorithm had a
  $\mathcal{O}(n \log n)$ complexity, caching can reduce it to $\mathcal{O}(n)$
  \cite{somenzi1998cudd}. Either way, the complexity of projection is still
  $\mathcal{O}(n^2)$.
\end{proof}

% NOTE: this is a rehash of equation (2) from the DPMC paper
\begin{definition} \label{def:abstraction}
  Let $(T, r, \gamma, \pi)$ be a PJT, and $X$ be the set of variables. The
  functionality of \textsc{DPMC} execution can be represented by $\delta(r)$,
  where $\delta\colon \mathcal{V}(T) \to \mathbb{R}^{2^X}$ is a recursive
  function defined as
  \begin{equation} \label{eq:execution}
    \delta(t) \coloneqq
    \begin{cases}
      \gamma(t) & \text{if } t \in \mathcal{L}(T) \\
      \exists_{\pi(t)} \prod_{u \in \mathcal{C}(t)} \delta(u) & \text{otherwise.}
    \end{cases}
  \end{equation}
  The range of $\delta(r)$ then contains a single real number, i.e., the answer.
\end{definition}

% TODO: will need to rephrase this to fit my context
\begin{theorem}[Theorem~4 in \cite{DBLP:conf/cp/DudekPV20}, almost
  verbatim] \label{thm:connection}
  Let $\phi$ be a propositional formula in CNF over a set $X$ of variables and
  $(S, \chi)$ be a tree decomposition of the primal graph of $\phi$ of width
  $w$. Then Algorithm~2 returns a PJT of $\phi$ of width at most $w+1$.
\end{theorem}

>>>>

\documentclass[runningheads]{llncs}

\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage[capitalise]{cleveref}

@@

@@

%\title{Contribution Title\thanks{Supported by organization x.}}
\title{Weighted Model Counting Without Parameter Variables}

%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here

%\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
%Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}

%\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.

%\institute{Princeton University, Princeton NJ 08544, USA \and
%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
%\email{lncs@springer.com}\\
%\url{http://www.springer.com/gp/computer-science/lncs} \and
%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
%\email{\{abc,lncs\}@uni-heidelberg.de}}

\maketitle              % typeset the header of the contribution

\begin{abstract}
  The abstract should briefly summarize the contents of the paper in
  150--250 words.

  \keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}

@@

% NOTE: ...that we will need.
% NOTE: The proof immediately comes from the definition (of operations).
\begin{proposition}[Basic properties] \label{prop:basic}
  For any propositional formulas $\phi$ and $\psi$, and $a, b, c, d \in
  \mathbb{R}$,
  \begin{itemize}
  \item $[\phi]^a_b = [\neg \phi]^b_a$;
  \item $c + [\phi]^a_b = [\phi]^{a+c}_{b+c}$;
  \item $c \cdot [\phi]^a_b = [\phi]^{ac}_{bc}$;
  \item $[\phi]^a_b \cdot [\phi]^c_d = [\phi]^{ac}_{bd}$;
  \item $[\phi]^1_0 \cdot [\psi]_0^1 = [\phi \land \psi]_0^1$.
  \end{itemize}
  And for any pair of pseudo-Boolean functions $f, g \colon 2^X \to \mathbb{R}$
  and $x \in  X$, $(fg)|_{x=i} = f|_{x=i} \cdot g|_{x=i}$ for $i = 0, 1$.
\end{proposition}

@@

\paragraph{Notes.}
\begin{itemize}
\item Apparently, the \textsc{DPMC} paper already shows that taking the first
  offered decomposition tree is best.
\item It is already well-known that WMC is FPT.
\item Weights on literals other than the positive literals that correspond to
  variables in $X_P$ are redundant as they either are equal to one or duplicate
  an already-defined weight.
\item Mention that formulas, clauses, and models are all treated as sets.
\item The constant is inspired by the \texttt{bklm16} encoding.
\item We don't make a distinction between a real number and a (pseudo-Boolean)
  function that always returns that number.
\item Throughout the paper, projection is assumed to have the lowest precedence
  (e.g., $\exists_x fg = \exists_x (fg)$ and not $(\exists_xf)g$).
\end{itemize}

@@

\begin{algorithm}
  \caption{WMC instance transformation}
  \label{alg:transformation}
  \SetKwFunction{rename}{rename}
  \SetKwProg{Fn}{Function}{:}{}
  \KwData{an (old-format) WMC instance $(\phi, X_I, X_P, w)$}
  \KwResult{a (new-format) WMC instance $(F, X, \omega)$}
  $F \gets \emptyset$\;
  $\omega \gets 1$\;
  let $d\colon X_P \to \mathbb{N}$ be defined as $p \mapsto |\{ o \in X_P \mid o
  \le p \}|$\;
  \ForEach{clause $c \in \phi$\label{line:foreach1start}}{
    \uIf{$c \cap X_P = \{ p \}$ for some $p$ \textnormal{\textbf{and}} $w(p) \ne
      1$}{
      \uIf{$|c| = 1$}{$\omega \gets \omega \times w(p)$\;}
      \Else{
        $F \gets F \cup \left\{ \left[ \bigwedge_{l \in c \setminus
              \{p\}} \neg l \right]^{w(p)}_1 \right\}$\;
      }
    }
    \ElseIf{$\{p \mid \neg p \in c\} \cap X_P  = \emptyset$}{
      $F \gets F \cup \{ [c]^1_0 \}$\; \label{line:foreach1end}
    }
  }
  \ForEach{indicator variable $v \in X_I$ \label{line:foreach2start}}{
    \If{$\{[v]_1^p, [\neg v]_1^q\} \subseteq F$ for some $p$ and $q$}{
      $F \gets F \setminus \{ [v]_1^p, [\neg v]_1^q \} \cup \{ [v]_q^p
      \}$\; \label{line:foreach2end}
    }
  }
  replace every variable $v$ in $F$ with $\rename{$v$}$\;
  \Return{$(F, X_I, \omega)$}\;
  \Fn{\rename{$v$}}{
    $S \gets \{u \in X_P \mid u \le v\}$\;
    \lIf{$S = \emptyset$}{\Return{$v$}}
    \Return{$v - d(\max S)$}\;
  }
\end{algorithm}

@@

\begin{theorem}[Correctness]
  \Cref{alg:transformation}, when given an old WMC instance $(\phi, X_I, X_P,
  w)$, returns a new WMC instance with the same answer, provided the following
  conditions are satisfied:
  \begin{enumerate}
  \item for all indicator variables $i \in X_I$, $w(i) = w(\neg i) = 1$,
  \item and either
    \begin{enumerate}
    \item for all parameter variables $p \in X_P$, there is a non-empty family
      of literals $(l_i)_{i=1}^n$ such that \label{condition:d02}
      \begin{enumerate}
      \item $w(\neg p) = 1$,
      \item $l_i \in X_I$ or $\neg l_i \in X_I$ for all $i = 1, \dots, n$,
        \label{condition:equivalence1}
      \item and $\{ c \in \phi \mid p \in c \text{ or } \neg p \in c \} =
        \left\{p \lor \bigvee_{i=1}^n \neg l_i \right\} \cup \{ l_i \lor \neg p
        \mid i = 1, \dots, n \}$; \label{condition:equivalence2}
      \end{enumerate}
    \item or for all parameter variables $p \in X_P$, \label{condition:bklm16}
      \begin{enumerate}
      \item $w(p) + w(\neg p) = 1$,
      \item for any clause $c \in \phi$, $|c \cap X_P| \le 1$,
      \item there is no clause $c \in \phi$ such that $\neg p \in c$,
      \item if $\{ p \} \in \phi$, then there is no clause $c \in \phi$ such
        that $c \ne \{ p \}$ and $p \in c$, \label{cond:just_parameter}
        % 'prime implicants are logically disjoint'
      \item and for any $c, d \in \phi$ such that $c \ne d$, $p \in c$ and $p
        \in d$, $\bigwedge_{l \in c \setminus \{ p \}} \neg l \land \bigwedge_{l
          \in d \setminus \{ p \}} \neg l$ is false. \label{cond:disjoint}
      \end{enumerate}
    \end{enumerate}
  \end{enumerate}
\end{theorem}
\begin{proof}
  By \cref{prop:equivalence},
  \begin{equation}
    \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
          x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
      1\right) \label{eq:new_wmc}
  \end{equation}
  is a new WMC instance with the same answer as the given old WMC instance. By
  \cref{def:new_wmc}, its answer is
  \begin{equation} \label{eq:answer}
    \left(\exists_{X_I \cup X_P} \prod_{c \in \phi} [c]_0^1 \prod_{x \in X_I
        \cup X_P} [x]_{w(\neg x)}^{w(x)} \right)(\emptyset)
  \end{equation}
  Since both \cref{condition:d02,condition:bklm16} ensure that each clause in
  $\phi$ has at most one parameter variable, we can partition $\phi$ into
  $\phi_* \coloneqq \{c \in \phi \mid \mathtt{Vars}(c) \cap X_P = \emptyset \}$
  and $\phi_p \coloneqq \{ c \in \phi \mid \mathtt{Vars}(c) \cap X_P = \{ p \}
  \}$ for all $p \in X_P$. We can then use \cref{thm:early} to reorder
  \cref{eq:answer} into
  \[
    \left(\exists_{X_I} \left( \prod_{x \in X_I} [x]_{w(\neg x)}^{w(x)} \right)
      \left( \prod_{c \in \phi_*} [c]_0^1 \right) \prod_{p \in X_P} \exists_p
      [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 \right)(\emptyset).
  \]

@@

  The only remaining case is when $w(p) \ne 1$ and $\{ p \} \not \in \phi_p$.
  Then $\omega = 1$, and
  \[
    F_p = \left\{ \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
        l\right]_1^{w(p)} \;\middle|\; c \in \phi_p \right\},
  \]
  so need to show that
  \[
    \prod_{c \in \phi_p} \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
      l\right]_1^{w(p)} = \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p}
    [c]_0^1.
  \]
  We can rearrange the right-hand-side as
  \begin{align*}
    \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[\bigwedge_{c \in \phi_p} c\right]_0^1 = \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 = w(p) + \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^{1-w(p)} \\
                                                               &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \neg \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \neg(c \setminus \{ p \}) \right]_1^{w(p)} \\
                                                               &= \left[ \bigvee_{c \in \phi_p} \neg \bigvee_{l \in c \setminus \{ p \}} l \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}
  \end{align*}
   by \cref{prop:basic,def:projection}. By \cref{cond:disjoint}, $\bigwedge_{l
     \in c \setminus \{ p \}} \neg l$ can be true for at most one $c \in
   \phi_p$, and so
   \[
     \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l
     \right]_1^{w(p)} = \prod_{c \in \phi_p} \left[ \bigwedge_{l \in c \setminus
         \{ p \}} \neg l \right]_1^{w(p)}
   \]
   which is exactly what we needed to show. This ends the proof that the
   first loop of \cref{alg:transformation} preserves the answer under both
   \cref{condition:d02} and \cref{condition:bklm16}. Finally, the loop on
   \crefrange{line:foreach2start}{line:foreach2end} of \cref{alg:transformation}
   replaces $[v]_1^p[\neg v]_1^q$ with $[v]_q^p$ (for some $v \in X_I$ and $p, q
   \in \mathbb{R}$), but, of course,
   \[
     [v]_1^p[\neg v]_1^q = [v]_1^p[v]_q^1 = [v]_q^p
   \]
   by \cref{prop:basic}.
\end{proof}

\paragraph{Notes.}
\begin{itemize}
\item This just says that $p$ is equivalent to a conjunction.
\item The first group of conditions applies to \texttt{d02}, while the second
  group applies to \texttt{bklm16}.
\item For \texttt{cd05} and \texttt{cd06}, condition $2bi$ should be replaced
  with $w(\neg p) = 1$.
\item Benefits of having this proof in the paper:
  \begin{itemize}
  \item It puts all encodings on a common ground.
  \item It illustrates the convenience of our notation for reasoning about
    (certain types of) pseudo-Boolean functions.
  \item It's too big and too important to be left for the appendix.
  \end{itemize}
\item Processors:
  \begin{itemize}
  \item Intel(R) Xeon(R) Gold 6138 CPU @ 2.00GHz
  \item Intel(R) Xeon(R) CPU E5-2630 v3 @ 2.40GHz
  \item Intel(R) Xeon(R) CPU E7-4820 v2 @ 2.00GHz
  \end{itemize}
\item Table captions should be placed above the tables.
\item Figures should preferably be in EPS format.
\item the environments 'theorem', 'definition', 'lemma', 'proposition',
  'corollary', 'remark', and 'example' are defined in the LLNCS documentclass as
  well.
\end{itemize}

\paragraph{TODO}
\begin{itemize}
\item Come up with better names for old/new WMC instances.
\item Do I need to formally consider extending a pseudo-Boolean function to a
  bigger domain?
\item Define scalar operations on ADDs.
\item Define projection of a set of variables.
\item add parentheses around products
\item maybe give a name to (2) since it's wrong to call it an equation
\item check if each condition is actually used. Maybe turn this into a paragraph
  that gives an overview of the proof.
\item condition 1 is necessary in both cases because we're ignore the weights of
  indicator variables
\item maybe do more summations over domains in proofs
\end{itemize}

\bibliographystyle{splncs04}
\bibliography{paper}
